Extended Abstract

### 2021-05-21 paper/paper.tex
Another pass

\begin{abstract}
  To type check a \picalc{} process with linear and shared types without relying on type annotations we need to infer the types of the channels we create within the process.
  If we were to limit ourselves to the shared \picalc{}, we could traverse a process bottom-up and always eagerly solve typing constraints into substitutions that are then applied to the typing context.
  However in our setting, typing constraints do not always come with have a most general unifier, and thus cannot always be eagerly solved.
  We follow Padovani \cite{Padovani15} and provide a \emph{co-contextual typing inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up and, in addition to the typing context, collecting typing constraints too.
  We then solve those constraints with a most general solution by using well-known unification algorithms \cite{McBride03}, and defer the rest until more information becomes available.

  We state clear soundness and completeness theorems for both these phases and a progress theorem that ensures that only those constraints without a most general unifier are deferred.
  The totality of this work is being mechanised in Agda.
\end{abstract}

@@

To type check a \picalc{} process with shared and linear types we must assign a type to every communication channel created within a process.
To do so we can either ask the user for type annotations \cite{ZalakainD21}, or we can synthesize types by looking at how channels are used.
We follow this latter approach by traversing processes bottom-up while keeping a typing context containing \emph{metavariables} (\emph{holes}) and collecting typing constraints on metavariables, in line with co-contextual typing inference \cite{ErdwegBKKM15}.
Keeping a strictly bottom-up information flow has the additional advantage of making typing inference easily parallelisable.
Constraints with a most general solution can safely be solved into substitutions and applied to the typing context, constraints without one must be kept around for later: substituting by a solution that is not the most general risks over-constraining the problem down the line.
Armed with a typing inference algorithm that for a process $P$ infers the most general typing context $\Gamma$ and some typing constraints, type checking that $\Delta \vdash P$ for some $\Delta$ (assuming the metavariables in $\Gamma$ and $\Delta$ are disjoint) amounts to emitting the extra constraint $\Gamma = \Delta$.

>>>>

\begin{abstract}
  To type check a \picalc{} process with linear and shared types without relying on type annotations we need to infer the types of the channels we create within a process.
  If we were to limit ourselves to the shared \picalc{}, we could traverse a process bottom-up and eagerly solve typing constraints into substitutions that are then applied to the typing context.
  However in our setting, typing constraints do not always come with have a most general solution, and thus cannot always be eagerly solved.

  We follow Padovani \cite{Padovani15} and provide a \emph{co-contextual typing inference} \cite{ErdwegBKKM15} algorithm that traverses processes bottom-up and, in addition to the typing context, collects a set of typing constraints.
  We then solve those constraints that have a most general solution by using well-known unification algorithms \cite{McBride03}, and defer the rest until more information becomes available.
  We state clear soundness and completeness theorems for both these phases and a progress theorem that ensures that only those constraints without a most general unifier are deferred.
  The totality of this work is being mechanised in Agda.
\end{abstract}

@@

To type check a \picalc{} process with shared and linear types we must assign a type to every communication channel created within a process.
To do so we can either ask the user for type annotations \cite{ZalakainD21}, or we can synthesize types by looking at how channels are used.
We follow this latter approach by traversing processes bottom-up while keeping a typing context containing \emph{metavariables} (\emph{holes}) and collecting typing constraints on metavariables, in line with co-contextual typing inference \cite{ErdwegBKKM15}.
Keeping a strictly bottom-up information flow has the additional advantage of making typing inference easily parallelisable.
Constraints with a most general solution can safely be solved into substitutions and applied to the typing context, constraints without one must be kept around for later: applying them as substitutions risks over-constraining the problem down the line.
Armed with a typing inference algorithm that for a process $P$ infers the most general typing context $\Gamma$ and some typing constraints, type checking that $\Delta \vdash P$ for some $\Delta$ amounts to emitting the extra constraint $\Gamma = \Delta$ (assuming the metavariables in $\Gamma$ and $\Delta$ are disjoint).
